---
title: "Magellan"
subtitle: "A developer's guidelines"
author: "Samuel Wieczorek"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    theme: united
    keep_md: true
    papersize: a4
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```


```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```


Magellan is a Shiny application which contains only few R6 classes. It is not used alone but it is the core engine for data processing workflows.
Thus, this article is not for end-users but for developers who are interesting in writing new data processing tools and/or to combine them in pipelines

# Introduction


# Global architecture


A process (or pipeline) consists in screens and it is a R6 class which inherits from classes in Magellan


# Writing new process

A process such as usable with Magellan is a R6 classes which contains:
* the configuration of the process,
* the list of screens with their severs (shiny)

This class inherits from classes in Magellan: this is how the core engine is integrated in xxx.


# Writing a new pipeline

