---
title: "Use Magellan with external R package"
subtitle: "A developer's guideline"
author: "Samuel Wieczorek"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  BiocStyle::html_document:
  toc: true
number_sections: true
theme: united
keep_md: true
papersize: a4
vignette: >
  %\VignetteIndexEntry{Use Magellan with external R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```


```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```



# Intro


The package `Magellan` implements a workflow manager which is generic at two points of view

First, it is suitable for objects of type list for which the fundamentla requirments are: 

* A list of items
* Each item is named and can be accessed either by its name or index
* The exists an 'append' function and a 'Remove' function. The first one adds an additionnal item at the end of the list (append function). The last one can remove any of the item in the list (even with a range)
We hope that the source code of modules in `Magellan` is sufficient to understand how it is structured. We give here several additional tips for a better understanding.


## External R package with modules

`Magellan` cannot be run alone, it is always run with a package which contains the informations about the steps to manage.

## Requirments

This package must have xxx.
Two other articles are available that explain how to write a module for simple and composed workflows.


# Specific functions to write

Whether the class of the object, it may be necessary to write the code for two functions that cannot be generic

* An append function named Add_Datasets_to_Object()
* A remove function named Keep_Datasets_from_Object()

Those functions must be part of the external package to avoid too many dependancies for the package `Magellan`.

The developer of shiny modules that will use `Magellan` can find in  an exmaple of those functions that are implemented for the type `list` and for examples purpose.

# Simple workflow


# Composed workflow